---
title: "p8105_hw3_zq2227.Rmd"
author: "Zixuan Qiu zq2227"
date: "2023-10-08"
output: github_document
---
# Problem 1
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

```{r}
aislesnumber<-instacart |>
  group_by(aisle) |># groups the instacart dataset by aisle.
  summarize(number = n())|>#count the number of order in each aisle.
  arrange(desc(number))
```
```{r}
aislesnumber
```
#There are 134 aisles and the fresh vegetables are the most items ordered from.
```{r}
aisleplot<-aislesnumber|>
  filter(number>10000)|>
  ggplot(aes(x = reorder(aisle, number), y = number)) + #arrange the aisle by the number of items
  geom_bar(stat = "identity")+
  coord_flip()+ # display the aisle name horizontal
  labs(
    title = "Number of items ordered in each aisle ",
    x = "Number of ordered",
    y = "Aisle",
    caption = "Data from the instcart")

aisleplot
ggsave("Item_number_in_aisle.png",plot =aisleplot )
```

```{r message=FALSE}
popular_items<-instacart|>
  filter(aisle %in% c("baking ingredients", "dog food care","packaged vegetables fruits"))|>
  group_by(aisle,product_name)|>
  summarize(ordered_number= n())|>
  arrange(aisle, desc(ordered_number))|>
  slice_head(n = 3)|>
  knitr::kable(digits = 1)
popular_items
```

```{r message=FALSE}
mean_hour<-instacart|>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream","packaged vegetables fruits"))|>
  group_by(product_name,order_dow)|>
  summarise(meanhour=mean(order_hour_of_day))|>
  pivot_wider( names_from = "order_dow", values_from = "meanhour")|>
  rename("Sunday" =`0`,"Monday" =`1`,"Tuesday" =`2`,"Wednesday" = `3`,"Thursday" =`4`,"Friday" =`5`,"Saturday" =`6`, "prdouct mean hour of the day"=product_name)|>
  knitr::kable(digits = 1)
mean_hour
```

# Problem 2 
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

```{r}
brffssdata=brfss_smart2010|>
  janitor::clean_names()|>
  filter(topic=="Overall Health")|>
  mutate(response=factor(response,order=TRUE,levels=c("Poor", "Fair", "Good", "Very good", "Excellent")))
str(pull(brffssdata,response))
```
### (a)
```{r}
states_2002 <-brffssdata |>
  filter(year==2002)|>
  group_by(locationabbr)|>
  summarise(nobservations=n())|>
  mutate(nobservations=nobservations/5)|>
  filter(nobservations>=7)
states_2002
```
#### CT,FL,MA,NC,NJ,PA  6 states were observed at 7 or more location in 2002.
```{r}
states_2010 <-brffssdata |>
  filter(year==2010)|>
  group_by(locationabbr)|>
  summarise(nobservations=n())|>
  mutate(nobservations=nobservations/5)|>
  filter(nobservations>=7)
states_2010
```
#### CA,CO,FL,MA,MD,NC,NE,NJ,NY,OH,PA,SC,TX,WA  14 states were observed at 7 or more location in 2010.

### (b)
```{r}
exresponse<-brffssdata|>
  filter(response=="Excellent")|>
  group_by(year,locationabbr)|>
  summarise(meanvalue=mean(data_value,na.rm = TRUE))|>
  rename(state=locationabbr)|>
  group_by(state)
exresponse
```
```{r}
exresponseplot=ggplot(exresponse,aes(x=year,y=meanvalue,group=state,color=state))+geom_line()+
  labs(
    title = "Excellent response mean value  over time",
    x = "Time(year)",
    y = "Mean(%)",
    color = "States",
    caption = "Data from the brfss_smart2010") 

exresponseplot
ggsave("exresponseplot.png",plot=exresponseplot)
```

### (c)
```{r}
nyresponse=brffssdata|>
  filter(locationabbr=="NY")|>
  filter(year == 2010 | year == 2006)|>
  select(year,locationabbr,locationdesc,response,data_value)|>
  group_by(year)|>
  mutate(locationdesc = str_replace(locationdesc, "NY - ", ""))|>
  mutate(locationdesc = str_replace(locationdesc, " County", ""))
```

```{r}
newyorkplot=ggplot(nyresponse, aes(x = locationdesc, y =data_value, fill=response)) +geom_bar(stat = "identity",position = "fill") +facet_wrap(~ year)+coord_flip()+
labs(
    title = "Distribution of responses among locations in NY",
    x = "% of response(Poor to Excellent) ",
    y = "Location",
    fill ="Level",
    caption = "brfss_smart2010")
  

newyorkplot
ggsave("NewYork plot.png",plot=newyorkplot)
```

# Problem 3
## Tidy participants
```{r message=FALSE}
participants= read_csv(file = "./nhanes_covar.csv",skip =4,n_max = 250)|>
   janitor::clean_names()|>
  mutate(sex,sex=ifelse(sex==1,"Male","Female"))|>
  mutate(sex,sex=as_factor(sex))|>
  filter(age>=21)|>
  drop_na()|>
  mutate(education=as.character(education))|>
  mutate(education = recode(education,  "1" = "Low","2" = "Medium","3" = "High"))|>
  mutate(education=factor(education, levels = c("Low", "Medium", "High"), order= TRUE))

participants
```



### Gender_education_distribution
```{r message=FALSE}
genderdata<-participants|>
  group_by(sex,education)|>
  summarise(number=n())|>
  pivot_wider( names_from = "sex", values_from = "number")|>
  knitr::kable(digits = 1)

genderdata
```

### Age distribution
```{r}
agedistribute<-participants|>
  select(sex,education,age)
 
```
```{r}
agedistributeplot<-
  ggplot(agedistribute,aes(x=age,fill=sex))+ geom_density(alpha=0.2)+facet_grid(~ education)+viridis::scale_fill_viridis( name = "Gender",discrete = TRUE,option="plasma")+
labs(
    title = "Age distributions in each education category",
    x = "Age",
    y = "Density",
    caption = "MIMS")

agedistributeplot
ggsave("Age distribution.png",plot=agedistributeplot)
```
#### Comments agedistributeplot<-ggplot(agedistribute,aes(x=sex,y=age))+ geom_boxplot()+facet_grid(~ education)


## Tidy accelerometer data and merge 
```{r message=FALSE}
acceldata = read_csv(file = "./nhanes_accel.csv")|>
  janitor::clean_names()|>
  drop_na()
```
```{r}
mimsdata <- left_join(participants,acceldata,by= "seqn")
```
```{r}
 sumactivity<-mimsdata|>
  group_by(education,sex)|>
  pivot_longer(min1:min1440,names_to = "time",values_to = "activity")|>
  group_by(seqn)|>
  summarize(totalactivity = sum(activity))
```
```{r}
summimsdata<-
  inner_join(mimsdata,sumactivity,by= "seqn")|>
  relocate(seqn,sex,age,education,totalactivity,bmi)
summimsdata
```


### Total activities (y-axis) against age (x-axis)
```{r message=FALSE}
activity_ageplot=summimsdata|>
  group_by(education)|>
  ggplot(aes(x=age,y=totalactivity,color=sex))+
  geom_point(size=2.5)+
  facet_grid(~education)+
  geom_smooth(se = FALSE)+
  theme(legend.position = "bottom")+
  viridis::scale_color_viridis( name = "Gender",discrete = TRUE)+
   ggthemes::theme_excel()+
  labs(
    title = "Total activities VS age",
    x = "Age",
    y = "Totoal MIMS values",
    caption = "Data from the MIMS")


activity_ageplot
ggsave("activity_age.png",plot=activity_ageplot)
```
Comment on your plot.

### Three-panel plot that shows the 24-hour activity time 
```{r}
wholedayactivity=summimsdata|>
  group_by(education,sex)|>
  pivot_longer(min1:min1440,names_to = "time",values_to = "activity")|>
  group_by(seqn)
wholedayactivity  
```

```{r message=FALSE}
threepanelplot<-ggplot(wholedayactivity,aes(x=time,y=activity,color=sex))+geom_point()+facet_wrap(~ education, nrow =3)+geom_smooth(se = FALSE)+
  labs(
    title = "24-hour activity",
    x = "Time",
    y = "Activity",
    color = "Gender",
    caption = "MIMS")


threepanelplot
ggsave("24hour_Three-panel.png",plot=threepanelplot)
```


