---
title: "p8105_hw3_zq2227.Rmd"
author: "Zixuan Qiu zq2227"
date: "2023-10-08"
output: github_document
---
# Problem 1
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

```{r}
aislesnumber<-instacart |>
  group_by(aisle) |># groups the instacart dataset by aisle.
  summarize(number = n())|>#count the number of order in each aisle.
  arrange(desc(number))
```
```{r}
aislesnumber
```
#There are 134 aisles and the fresh vegetables are the most items ordered from.
```{r}
aisleplot<-aislesnumber|>
  filter(number>10000)|>
  ggplot(aes(x = reorder(aisle, number), y = number)) + #arrange the aisle by the number of items
  geom_bar(stat = "identity")+
  coord_flip()+ # display the aisle name horizontal
  labs(
    title = "Number of items ordered in each aisle ",
    x = "Aisle",
    y = "Number of ordered",
    caption = "Data from the instcart")

aisleplot
ggsave("Item_number_in_aisle.png",plot =aisleplot )
```

```{r message=FALSE}
popular_items<-instacart|>
  filter(aisle %in% c("baking ingredients", "dog food care","packaged vegetables fruits"))|>
  group_by(aisle,product_name)|>
  summarize(ordered_number= n())|>
  arrange(aisle, desc(ordered_number))|>
  slice_head(n = 3)|>
  knitr::kable(digits = 1)
popular_items
```

```{r message=FALSE}
mean_hour<-instacart|>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream","packaged vegetables fruits"))|>
  group_by(product_name,order_dow)|>
  summarise(meanhour=mean(order_hour_of_day))|>
  pivot_wider( names_from = "order_dow", values_from = "meanhour")|>
  rename("Sunday" =`0`,"Monday" =`1`,"Tuesday" =`2`,"Wednesday" = `3`,"Thursday" =`4`,"Friday" =`5`,"Saturday" =`6`, "prdouct mean hour of the day"=product_name)|>
  knitr::kable(digits = 1)
mean_hour
```

# Problem 2 
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

```{r}
brffssdata=brfss_smart2010|>
  janitor::clean_names()|>
  filter(topic=="Overall Health")|> #focus on the “Overall Health”  
  mutate(response=factor(response,order=TRUE,levels=c("Poor", "Fair", "Good", "Very good", "Excellent"))) # responses from “Excellent” to “Poor” factor from 5-1
str(pull(brffssdata,response))
```
### (a)
```{r}
states_2002 <-brffssdata |>
  filter(year==2002)|>
  group_by(locationabbr)|>
  summarise(nobservations=n())|>
  mutate(nobservations=nobservations/5)|>
  filter(nobservations>=7) #7 or more locations
states_2002
```
#### CT,FL,MA,NC,NJ,PA  6 states were observed at 7 or more location in 2002.
```{r}
states_2010 <-brffssdata |>
  filter(year==2010)|>
  group_by(locationabbr)|>
  summarise(nobservations=n())|>
  mutate(nobservations=nobservations/5)|>
  filter(nobservations>=7) # 7 or more locations
states_2010
```
#### CA,CO,FL,MA,MD,NC,NE,NJ,NY,OH,PA,SC,TX,WA  14 states were observed at 7 or more location in 2010.

### (b )
```{r message=FALSE}
exresponse<-brffssdata|>
  filter(response=="Excellent")|> #limited to Excellent responses
  group_by(year,locationabbr)|>
  summarise(meanvalue=mean(data_value,na.rm = TRUE))|> #averages  of data_value across locations
  rename(state=locationabbr)|>
  group_by(state)
exresponse
```

```{r}
exresponseplot=ggplot(exresponse,aes(x=year,y=meanvalue,group=state,color=state))+geom_line()+
  labs(
    title = "Excellent response mean value  over time",
    x = "Time(year)",
    y = "Mean(%)",
    color = "States",
    caption = "Data from the brfss_smart2010") 

exresponseplot
ggsave("exresponseplot.png",plot=exresponseplot)
```

### (c)
```{r}
nyresponse=brffssdata|>
  filter(locationabbr=="NY")|> #limited to NY
  filter(year == 2010 | year == 2006)|> # limited to 2006 and 2010
  select(year,locationabbr,locationdesc,response,data_value)|> 
  group_by(year)|>
  mutate(locationdesc = str_replace(locationdesc, "NY - ", ""))|># remove the  variables prefix 
  mutate(locationdesc = str_replace(locationdesc, " County", ""))# change variables name 
```

```{r}
newyorkplot=ggplot(nyresponse, aes(x = locationdesc, y =data_value, fill=response)) +geom_bar(stat = "identity") +facet_wrap(~ year)+coord_flip()+
labs(
    title = "Distribution of responses among locations in NY",
    x = "Location",
    y = "% of response(Poor to Excellent)",
    fill ="Level",
    caption = "brfss_smart2010")
  

newyorkplot
ggsave("NewYork plot.png",plot=newyorkplot)
```

# Problem 3
## Tidy participants
```{r message=FALSE}
participants= read_csv(file = "./nhanes_covar.csv",skip =4,n_max = 250)|>
   janitor::clean_names()|>
  mutate(sex,sex=ifelse(sex==1,"Male","Female"))|>
  mutate(sex,sex=as_factor(sex))|>
  filter(age>=21)|> #exclude participants less than 21 years of age
  drop_na()|># remove missing demographic data; 
  mutate(education=as.character(education))|>
  mutate(education = recode(education,  "1" = "Low","2" = "Medium","3" = "High"))|>
  mutate(education=factor(education, levels = c("Low", "Medium", "High"), order= TRUE)) # encode data with reasonable variable classes 

participants
```

### Gender_education_distribution
```{r message=FALSE}
genderdata<-participants|>
  group_by(sex,education)|>
  summarise(number=n())|>
  pivot_wider( names_from = "sex", values_from = "number")|>
  knitr::kable(digits = 1)

genderdata
```
#### Comment: There are total 228 participants in the study and 118 are male, 110 are female. In high education catagory, female is more than male. In medium education level, male has 12 more  participants than female.

### Age distribution
```{r}
agedistribute<-participants|>
  select(sex,education,age)
 
```
```{r}
agedistributeplot<-
  ggplot(agedistribute,aes(x=age,fill=sex))+ geom_density(alpha=0.2)+
  facet_grid(~ education)+
  viridis::scale_fill_viridis( name = "Gender",discrete = TRUE,option="plasma")+
  labs(
    title = "Age distributions in each education category",
    x = "Age",
    y = "Density",
    caption = "MIMS data")

agedistributeplot
ggsave("Age distribution.png",plot=agedistributeplot)
```
#### Comment:from the plot, the age distribution of high education level is much younger than the low and medium education level. And the high education felmale  whose age colse to  30 has a higher density than male. In the medium education level, the age distribution of the male were younger than female. And the low education level age distribute  is older than other two categorys.  



## Tidy accelerometer data and merge 
```{r message=FALSE}
acceldata = read_csv(file = "./nhanes_accel.csv")|>
  janitor::clean_names()|>
  drop_na()
```
```{r}
mimsdata <- left_join(participants,acceldata,by= "seqn")
```
```{r}
 sumactivity<-mimsdata|>
  group_by(education,sex)|>
  pivot_longer(min1:min1440,names_to = "time",values_to = "activity")|>
  group_by(seqn)|>
  summarize(totalactivity = sum(activity))# aggregate across minutes to create a total activity df
```
```{r}
summimsdata<-
  inner_join(mimsdata,sumactivity,by= "seqn")|> # bind the  total activity variable into the mims  data frame 
  relocate(seqn,sex,age,education,totalactivity,bmi)
summimsdata
```


### Total activities (y-axis) against age (x-axis)
```{r message=FALSE}
activity_ageplot=summimsdata|>
  group_by(education)|>
  ggplot(aes(x=age,y=totalactivity,color=sex))+
  geom_point(size=3)+
  facet_grid(~education)+
  geom_smooth(method = "lm",se = FALSE)+
  theme(legend.position = "bottom")+
  viridis::scale_color_viridis( name = "Gender",discrete = TRUE)+
   ggthemes::theme_excel()+ #change to eccel theme 
  labs(
    title = "Total activities VS Age",
    x = "Age",
    y = "Totoal MIMS values",
    caption = "Data from MIMS")


activity_ageplot
ggsave("activity_age.png",plot=activity_ageplot)
```
#### Comment: In medium and high education level,the female always has more activity than male who has same age. And the young people has more activity than older people especially in low education level.  The high education old people has more activity than other education level  but the young people in high education has less activity compare to other level in same age, in other words, the total activity-age distribution  of the high education is more concentrated or flat.

### Three-panel plot that shows the 24-hour activity time 
```{r}
wholedayactivity=summimsdata|>
  group_by(education,sex)|>
  pivot_longer(min1:min1440,names_to = "time",values_to = "activity",names_prefix = "min")|>
  mutate(time=as.numeric(time))|>
  group_by(time,education,sex)|>
  mutate(meanactivity =mean(activity))

wholedayactivity  
```

```{r message=FALSE}
threepanelplot<-ggplot(wholedayactivity,aes(x=time,y=meanactivity,color=sex))+
  geom_point()+
  facet_wrap(~ education, nrow =3)+
  geom_smooth(method="lm", se=FALSE)+
  labs(
    title = "24-hour activity",
    x = "Time (24hour)",
    y = "Activity",
    color = "Gender",
    caption = "MIMS")

threepanelplot
ggsave("24hour_Three-panel.png",plot=threepanelplot)
```
#### Comment：the ctivity strnegth is start at midnight with a low valuse , decrease to a clear trough, then a gradual rise, a peak, and then a decline. This meets a  general pattern of daily activity, such as sleep at night(0minto 300min,12am to 6am), starting activity in the morning, peaking in the noon(700min=12.pm), and then decreasing as the day transitions into the evening.
#### In high education level, the difference between genders is more noticeable, with females consistently having higher activity levels than males throughout the day.
#### In low education level, the peek activity strength is higher than other education level and there are no significant activity difference between male and female. 

